Primer on Production Functions and Economic Growth

### 2014-11-28 Articles/Prod Func Primer/ProdFuncPrimer.tex
Minor edits to Primer.

\usepackage{wrapfig}    % Allows wrapping of text around figures
\usepackage{float}      % Allows precise positioning of tables and figures within the text
\usepackage{amsmath}    % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow}   % To create multirow tables
\usepackage{hyperref}   % To create hyperlinks in the paper
\usepackage{microtype}  % Produces hanging punctuation and beautiful type
\usepackage{booktabs}   % For nice tables

@@

The LinEx model does, however, include terms that represent
efficiencies among capital, labor, and energy, in the form of 
the ratios $\rho_k$ and $\rho_l$ which are defined by
%
\begin{equation} \label{eq:rho_k}
  \rho_k \equiv \frac{k}{(1/2)(l+e)}
\end{equation}
%
and
%
\begin{equation} \label{eq:rho_l}
  \rho_l \equiv \frac{l}{e}
\end{equation}
%
and represent 
%
(\ref{eq:rho_k}) capital stock increase relative to the average of labor and energy (capital deepening) and 
(\ref{eq:rho_l}) labor increase relative to energy.
The effect of energy ($e$) on output ($y$) is enhanced if either
%
\begin{itemize}
\item{capital stock ($k$) has increased more than the average 
      of labor ($l$) and energy ($e$), i.e. if $\rho_k >$ 1, or}
\item{labor ($l$) has increased more than energy ($e$), i.e. if $\rho_l >$ 1.}
\end{itemize}
%
An economy that increases 
capital stock ($k$) and labor ($l$) at the same rate as energy will have
capital and labor efficiency ratios ($\rho_k$ and $\rho_l$, respectively) of unity.
In that case, economic output ($y$) increases at the same rate as energy consumption ($e$).
According to the LiNex model, an economy that increases 
capital stock ($k$) without a commensurate increase in labor ($l$) 
and energy ($e$) will experience an increase in output ($y$) in excess 
of its increase of energy consumption ($e$), because
%
\begin{equation}
    \mathrm{e}^{a \left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]}
	> 1 \; .
\end{equation}
%
Similarly, an economy benefits by increasing labor ($l$) 
without a commensurate increase in energy ($e$). 
Thus, $\rho_k$ is an 
efficiency of using additional labor and energy to make 
additional capital stock available to the economy, 
and $\rho_l$ is an efficiency of using additional energy to make 
additional labor available to the economy. 
Increases in $\rho_k$ and $\rho_l$ provide upward pressure on economic output ($y$),
as the only factor of production ($e$) is used more efficiently.

The parameter $a$ represents the overall importance of efficiencies $\rho_k$ and $\rho_l$, 
and the parameter $c$ represents the relative importance 
of the labor efficiency term ($\rho_l$) compared to the capital stock efficiency term ($\rho_k$).

\citet{Kummel:2010vz} use primary exergy for $e$ and find that 
both $a$ and $c$ vary with time.
\citet{Warr:2012cg} use useful work for $u$ and find that 
constant values of $a$ and $c$ are sufficient to describe economic growth 
over the long run.

>>>>

\usepackage{wrapfig}    % Allows wrapping of text around figures
\usepackage{float}      % Allows precise positioning of tables and figures within the text
\usepackage{amsmath}    % Allows \begin{equation*} \end{equation*} for unnumbered equations
\usepackage{multirow}   % To create multirow tables
\usepackage{hyperref}   % To create hyperlinks in the paper
\usepackage{microtype}  % Produces hanging punctuation and beautiful type
\usepackage{booktabs}   % For nice tables
\usepackage{nicefrac}   % For beautiful fractions

@@

The LinEx model does, however, include terms that represent
efficiencies among capital, labor, and energy, in the form of 
the ratios $\rho_k$ and $\rho_l$ which are defined by
%
\begin{equation} \label{eq:rho_k}
  \rho_k \equiv \frac{k}{(\nicefrac{1}{2})(l+e)}
\end{equation}
%
and
%
\begin{equation} \label{eq:rho_l}
  \rho_l \equiv \frac{l}{e}
\end{equation}
%
and represent 
%
(\ref{eq:rho_k}) capital stock increase relative to the average of labor and energy (capital deepening) and 
(\ref{eq:rho_l}) labor increase relative to energy.
The effect of energy ($e$) on output ($y$) is enhanced if either
%
\begin{itemize}
\item{capital stock ($k$) has increased more than the average 
      of labor ($l$) and energy ($e$), i.e. if $\rho_k >$ 1, or}
\item{labor ($l$) has increased more than energy ($e$), i.e. if $\rho_l >$ 1.}
\end{itemize}
%
An economy that increases 
capital stock ($k$) and labor ($l$) at the same rate as energy ($e$) will have
capital and labor efficiency ratios ($\rho_k$ and $\rho_l$, respectively) of unity.
In that case, economic output ($y$) increases at the same rate as energy consumption ($e$).
According to the LiNex model, an economy that increases 
capital stock ($k$) without a commensurate increase in the average of labor
and energy [$(\nicefrac{1}{2})(l+e)$] will experience an increase in output ($y$) in excess 
of its increase of energy consumption ($e$), because $\rho_k > 1$ and
%
\begin{equation}
    \mathrm{e}^{a \left[2 \left(1 - \frac{1}{\rho_k} \right) 
    + c \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]}
	> 1 \; .
\end{equation}
%
Similarly, an economy benefits by increasing labor ($l$) 
without a commensurate increase in energy ($e$), because $\rho_l > 0$. 
Thus, $\rho_k$ is an 
efficiency of using additional labor~($l$) and energy~($e$) to make 
additional capital stock~($k$) available to the economy, 
and $\rho_l$ is an efficiency of using additional energy~($e$) to make 
additional labor~($l$) available to the economy. 
Increases in $\rho_k$ and $\rho_l$ provide upward pressure on economic output~($y$),
as the only factor of production~($e$) is used more efficiently.

The parameter $a$ represents the overall importance of efficiencies $\rho_k$ and $\rho_l$, 
and the parameter $c$ represents the relative importance 
of the labor efficiency term~($\rho_l$) compared to the capital stock efficiency term~($\rho_k$).

\citet{Kummel:2010vz} use primary exergy for $e$ and find that 
both $a$ and $c$ vary with time.
\citet{Warr:2012cg} use useful work for $e$ and find that 
constant values of $a$ and $c$ are sufficient to describe economic growth 
over the long run.
